---
title: "Answers_Assignment_01"
author: "lkoster"
date: "April 7, 2016"
output: html_document
---

library(dplyr)
library(broom)
library(texreg)

## Problem 3: Predicting Sprint Times

|:---|:---|
| `year` | Year of the Olympics or World Championships |
| `time` | Winning time |
| `women` | 1 if women's race; 0 if men's race |
| `olympics` | 1 if in the olympics; 0 if in the World Championships |

Load the data into R,
```{r, eval = FALSE}
sprinters <- read.csv("sprinters.csv")
```


#a. The referenced paper only used data from the Olympics 2004 and before. Create a new dataset named #`sprinters_orig` with only those observations.

sprinters_orig <- sprinters %>%
  filter (year %in% )

  select(varnames) %>%

  



b. Run the regressions
    ```{r, eval = FALSE}
  model1 <-   lm(time ~ year + women, data = sprinters_orig)
  model2 <- lm(time ~ year * women, data = sprinters_orig)
  model3 <-   lm(time ~ year, data = filter(sprinters_orig, women == 1))
  model4 <-   lm(time ~ year, data = filter(sprinters_orig, women == 0))
    ```
    Interpret each regression. How are they similar or different in their 
    slopes?
  
c. Plot the results of these regressions.
d. Use the function `predict` to predict the times of men and women in the 2156 Olympics. Is this plausible? 
